---
title: "Novaglio Et Al. 2021, Journal of Applied Ecology"
author: "Camilla Novaglio"
date: "27/12/2020"
output: 
  html_document:
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# File description  

This file is to be considered as Supplementary Information for the paper 'Exploring trade-offs in mixed fisheries by integrating fleet dynamics into multispecies size-spectrum models' submitted to the Journal of Applied Ecology.  

This file loads the data and all the codes needed to run the calibrated size-spectrum and fleet dynamics model that represents the Australia’s South East Shark and Scalefish Fishery. It provides the data and codes to set up scenarios, run projections and calculate ecological and socio-economic indicators detailed in the paper. Finally, this file reproduces all modelling and data figures in the paper, specifically Fig 2 (baseline model's output), Fig 3 (final model's outputs), Fig 4 (Scenarios and model projections to 2040) and Fig 5 (change in indicators across scenarios and relative to the status quo). For more information, please contact camilla.novaglio@gmail.com  

## Load data and codes  

```{r codes, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

####### load data 
rm(list=ls())
load("/Users/camillan/R-projects/mixed_fisheries_tradeoff/data/Mixed_fisheries_tradeoff.RData") 

####### load libraries 
library(tidyverse)
library(devtools)
library(plyr) 
library(Rcpp)
library(reshape2)
library(inline)
library(ggplot2)
library(ggthemes)
library(vegan)
library(mizer) # using "_mizer_inner_project_loop" function from Mizer package for faster iteration of species one time step forward (Project.R, line 573)
library(patchwork)
library(cowplot)
library(ggpubr)
library(matrixStats)

####### load model 
setwd("/Users/camillan/R-projects/mixed_fisheries_tradeoff/R/")
source("help.R") 
source("MizerParams-class.R") 
source("MizerSim-class.R")
source("plots.R")
source("project_methods.R")
source("project.R") 
source("RcppExports.R") 
source("selectivity_funcs.R")
source("summary_methods.R")
source("wrapper_functions.R")
source("DataFunction.R") 

####### specify or rename model inputs for the calibrated fleet dynamics and size-spectrum model
df_param = df_param2 # species-specific parameters 
interaction = theta # predator-prey interaction matrix
kappa = kappa # intercept of the pelagic background spectrum 
kappa_ben = kappa_ben # intercept of the benthic background spectrum # NOTE CAMI benthic is larger than plankton spectrum... 
kappa_alg = kappa_alg # intercept of the algae background spectrum (not used)
w_pp_cutoff = w_pp_cutoff # cut off size of the benthic background spectrum 
min_w_bb = min_w_bb # minimum size of the pelagic background spectrum 
w_bb_cutoff = w_bb_cutoff # cut off size of the benthic background spectrum 
fleetDynamics = TRUE
management = TRUE 
multiFleet = FALSE # TRUE only when fleets are driven by fleet-specific effort and not price and costs
selectivity_params = df_selParam_new # selectivity parameters 
catchability = df_target_bmsy2 # species-fleet catchability matrix (not considered as part of 'target' below)
target = df_target_bmsy2 # species-fleet interaction matrix
effort = 0 # fishing effort (not considered as emergent from the model)
dt = 0.25 # model time resolution
price = df_price_new # species market price  
cost = df_cost3 # fleet costs (already scaled)
# scaling_costs = scaling_costs # influential - set at a different value for each fleet
diet_steps = 0 # needed only to plot diet matrix 
ke = ke_fleet # fleets' time scale parameter 
scaling_price = scaling_price # not influential - set to 1
Blevel_management = "Bmsy" # management type 
initial_n = initial_n # initial species abundance 
initial_n_pp = initial_n_pp # initial pelagic background abundance 
initial_n_bb = initial_n_bb # initial benthic background abundance 
initial_effort = initial_effort2 # initial effort as a starting values for fleet dynamics 
scaling_effort<-rep(1,5) # not influential - set to 1

```

## Calibrated model

### Run calibrated baseline model and produce Figure 2

```{r run model baseline, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}

# sim_fitted is a Mizer object where all inputs to run the calibrated baseline model and all outputs from this model are already stored. 
# to run the model: 

# ####### set up model parameters 
# params_fitted <- MizerParams(sim_fitted@params@species_params, interaction = sim_fitted@params@interaction, kappa = sim_fitted@params@kappa, kappa_ben = sim_fitted@params@kappa_ben, kappa_alg = sim_fitted@params@kappa_alg, w_pp_cutoff = w_pp_cutoff, min_w_bb = min_w_bb, w_bb_cutoff = w_bb_cutoff, fleetDynamics = FALSE, selectivity_params = NA, catchability = NA, target = NA)
# 
# ####### run model
# sim_fitted<- project(params_fitted, effort = sim_fitted@effort, dt = dt, fleetDynamics = FALSE, multiFleet = FALSE, management = FALSE, price = NA, cost = NA, diet_steps = diet_steps)
# 
# ####### plots
# plot(sim_fitted)

###### Figure 2 
fig2<-compareTrends(sim_fitted,sim_FD = NA,fleetDynamics = FALSE,type = "yield",yieldObs_timeVariant,ssbObs = NA,rescale=3,areaEco)$plotYield_sp

# resize for publication 

theme_final<-theme(
  text = element_text(size=11, face = "plain"),
  legend.text=element_text(size=9),
  axis.title.y = element_text(size=11, vjust=2),
  axis.title.x = element_text(size=11, vjust=0.3),
  axis.text.y = element_text(size=9, hjust=0.95),
  axis.text.x = element_text(size=9, angle = 90, hjust=0.95))

fig2B<-fig2
fig2B<-fig2B+theme_final+ theme(strip.text.x = element_text(face = "plain"))

pdf("/Users/camillan/Dropbox/Mizer-fleet_extension/plot/FD/Final/revision/Final_uploaded/Figure_2.pdf", height=5, width=7.08)
fig2B
dev.off()

```

### Run calibrated fleet dynamics and size-spectrum model   

```{r run model, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}

####### set up model parameters 
params_FD <- MizerParams(df_param, interaction = interaction, kappa = kappa, kappa_ben = kappa_ben, kappa_alg = kappa_alg, w_pp_cutoff = w_pp_cutoff, min_w_bb = min_w_bb, w_bb_cutoff = w_bb_cutoff, fleetDynamics = fleetDynamics, selectivity_params = selectivity_params, catchability = catchability, target = target)

####### run model
sim_FD_bmsy <- project(params_FD, effort = effort, dt = dt, fleetDynamics = fleetDynamics, multiFleet = multiFleet, management = management, price = price, cost = cost, diet_steps = diet_steps, ke = ke, initial_n = initial_n, initial_n_pp = initial_n_pp, initial_n_bb = initial_n_bb, initial_effort = initial_effort, scaling_price = scaling_price, Blevel_management = Blevel_management) 

####### plots
plot_CN(sim_FD_bmsy)

```

### Figure 3   

```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}

####### Panel A - effort dynamics 

# extract modeled and observed data 
yieldData_fleet<-compareTrends(sim_fitted,sim_FD = sim_FD_bmsy, fleetDynamics = TRUE,type = "yield",yieldObs_timeVariant,ssbObs=NA,rescale=0,areaEco)$DataPlotYield_fl 
colnames(yieldData_fleet)<-c("year", "fleet","yield_tonnes","type")
effortData<-compareTrends(sim_fitted,sim_FD = sim_FD_bmsy, fleetDynamics = TRUE,type = "yield",yieldObs_timeVariant,ssbObs=NA,rescale=0,areaEco)$DataPlotEffort_fl 
colnames(effortData)<-c("year", "fleet","effort_opn","type")

# arrange data 
a<-effortData
b<-yieldData_fleet

colnames(a)<-c("year", "fleet", "EFFORT", "color")
colnames(b)<-c("year", "fleet", "YIELD", "color")
temp<-merge(a, b, all = TRUE)
temp<-temp %>% 
  gather("type", "value", -c(year, fleet, color))

a = c("SET-SH","SET-US","SSG","SET-DS","SED")
temp2<-temp %>% 
  spread(color, value) %>% 
  filter(type == "EFFORT") %>% 
  mutate(fleet = factor(fleet, level = a)) 

# plot data 
scaleFUN <- function(x) sprintf("%.0f", x)

plotEffortCatch_fl <- ggplot(temp2) + 
  geom_line(aes(x=year, y= modelled_FD), size=0.5) +
  geom_point(aes(x=year, y = observed), size = 1, shape = 1)+
  scale_x_continuous(labels = scaleFUN, name = "Year") +
  theme_tufte(base_family = "") +
  theme(text = element_text(size=14),
        axis.title.y = element_text(vjust=0.4, size = 15),
        axis.title.x = element_text(vjust=0.3, size = 15),
        axis.text.y = element_text(size=10, hjust=0.5),
        axis.text.x = element_text(size=10, angle = 90, hjust=0.5),
        panel.grid.major = element_blank(),
        axis.line.x = element_line(color="black"),
        axis.line.y = element_line(color="black"),
        panel.border = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_text(face = "bold"))+
  scale_y_continuous(name = "Effort (n. of hauls)")+
  facet_wrap(~fleet, nrow = 1)+
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)

# resize for publication 
plotEffortCatch_fl<-plotEffortCatch_fl +theme_final + theme(strip.text.x = element_text(face = "plain"))

####### Panel B - catch composition

# plot data using function in DataFunction.R 
fig3b<-catchComp(sim_FD_bmsy, datValidationYieldSppFl)$p

# resize for publication 
fig3b<-fig3b +theme_final+ theme(strip.text.x = element_text(face = "plain"))

# fix legends 
leg<-get_legend(fig3b)
leg<-as_ggplot(leg)
fig3b2<-fig3b + theme(legend.position = "none")

####### final figure 

# add captions 
plotEffortCatch_fl <- plotEffortCatch_fl + ggtitle("a)") + theme(plot.title = element_text(size = 11, face = "bold"))
fig3b2 <- fig3b2 + ggtitle("b)") + theme(plot.title = element_text(size = 11, face = "bold"))

fig3<-((plotEffortCatch_fl/fig3b2 + plot_layout(nrow=2,height=c(1,2))) | (plot_spacer()/leg + plot_layout(nrow=2,height=c(1,1)))) + plot_layout(ncol=2, widths = c(4,1))

pdf("/Users/camillan/Dropbox/Mizer-fleet_extension/plot/FD/Final/revision/Final_uploaded/Figure_3.pdf", height=6, width=7.08)
fig3
dev.off()

```

## Scenarios   

### Update model inputs  

```{r scenarios, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

#######  specify starting model for projections 
sim = sim_FD_bmsy

####### initial abundances as output of calibrated model (last year of model run)
initial_n_scenario = sim@n[nrow(sim@n),,]
initial_n_pp_scenario  = sim@n_pp[nrow(sim@n_pp),]
initial_n_bb_scenario  = sim@n_bb[nrow(sim@n_bb),]
initial_effort_scenario = sim@effortOut[dim(sim@effortOut)[1],]   

####### forward price and costs to 2040
t_max = 23 # use 83 to run the model to 2100 (Fig S9)
price_forward = t(replicate(t_max, price[nrow(price),]))
rownames(price_forward)<-seq(2018,(2018+t_max-1))

cost_forward1 <- t(replicate(t_max, cost[nrow(cost),,1]))
rownames(cost_forward1)<-seq(2018,(2018+t_max-1))

cost_forward2 <- t(replicate(t_max, cost[nrow(cost),,2]))
rownames(cost_forward2)<-seq(2018,(2018+t_max-1))

cost_forward <-array(c(cost_forward1,cost_forward2),dim = c(dim(cost_forward1)[1],dim(cost)[2],dim(cost)[3]), dimnames = c(dimnames(cost_forward1)[1],dimnames(cost)[2],dimnames(cost)[3]))

```

### Change fleet interactions

Here we define scenarios, each having a different fleet-species interaction matrix. This matrix determines each species' proportional contribution to the fleet's catch - hence this matrix defines competitive fleet interactions. The scenarios we consider are no competition, maximum competition, less bycatch, more valuable and more under-utilized. Below we compare these scenarios with the status quo (interactions as per catch and effort data).

```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

######## un-fished community
df_target_unfished<-target
df_target_unfished[]<-0

####### status quo 
# target

######## no competition
df_target_NoCompetition<-target
trial<-colMaxs(df_target_NoCompetition)
trial<-trial[-which(trial == 0)]
df_target_NoCompetition[df_target_NoCompetition %in% trial]<-1
df_target_NoCompetition[df_target_NoCompetition != 1]<-0

# keep Fmort constant across scenarios  
div<-colSums(df_target_NoCompetition*initial_effort_scenario)/colSums(target*initial_effort_scenario)
df_target_NoCompetition<-sweep(df_target_NoCompetition, 2, div, FUN = '/')
df_target_NoCompetition[, "myctophids"]<-0  

######## Max competition 
df_target_FullCompetition<-target
df_target_FullCompetition[]<-1
df_target_FullCompetition[, "myctophids"]<-0  

# keep Fmort constant across scenarios 
div<-colSums(df_target_FullCompetition*initial_effort_scenario)/colSums(target*initial_effort_scenario)
df_target_FullCompetition<-sweep(df_target_FullCompetition, 2, div, FUN = '/')
df_target_FullCompetition[, "myctophids"]<-0  

######## Less bycatch
df_target_NoBycatch<-target
# bycatch defined as: the 2 fleets with lowest proportion of catches should not catch the spp 
df<-target
for(i in 1:ncol(df)){
  df<-df[order(df[,i]),]
  df[c(1,2,3),i]<-0
}
df<-df[c(2,1,4,3,5),]
df_target_NoBycatch<-df

# keep Fmort constant across scenarios 
div<-colSums(df_target_NoBycatch*initial_effort_scenario)/colSums(target*initial_effort_scenario)
df_target_NoBycatch<-sweep(df_target_NoBycatch, 2, div, FUN = '/')
df_target_NoBycatch[, "myctophids"]<-0  

######## more valuable     
# define most valuable spp and increase fishing by 50% 
valuable<-sort(-df_price_new[nrow(df_price_new),])
valuable<-names(valuable)[1:6]

# increase Fmort of valuable by 50% of current Fmort and spread that increase across fleets proportionally
df_target_MoreTarget<-target
a<-colSums(target[,valuable]*initial_effort_scenario) 
a<-a + (a*0.5) 
div<-colSums(df_target_MoreTarget[,valuable]*initial_effort_scenario)/a
df_target_MoreTarget[,valuable]<-sweep(df_target_MoreTarget[,valuable], 2, div, FUN = '/')

######## target more under-quota   
# define under quota spp and increase fishing on these by 50%. https://www.afma.gov.au/sites/default/files/semac_tac_recommendations_2019-20.pdf
under<-c("macruronus novaezelandiae","nemadactylus macropterus","zeus faber","pristiophorus cirratus") 

# increase Fmort of under-quota by 50% of current Fmort and spread that increase across fleets proportionally
df_target_MoreUnderQuota<-target
a<-colSums(target[,under]*initial_effort_scenario) 
a<-a + (a*0.5)
div<-colSums(df_target_MoreUnderQuota[,under]*initial_effort_scenario)/a
df_target_MoreUnderQuota[,under]<-sweep(df_target_MoreUnderQuota[,under], 2, div, FUN = '/')

# Fishing mortality for each species from all fleets is <1 during all model and scenarios' runs - i.e. exploitation rate (catch/biomass at t) < 1

```

### Run scenarios 

Here we project the model forward (to 2040) for each scenarios 

```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

target_scenario = list(unfished = df_target_unfished, noCompetition = df_target_NoCompetition, fullCompetition = df_target_FullCompetition, statusQuo = target, noBycatch = df_target_NoBycatch, MoreTarget = df_target_MoreTarget, MoreUnderQuota = df_target_MoreUnderQuota) 

runModel<-function(target_scenario){
  
  params_scenario<-list()
  sim_scenario<-list()

  for(i in 1:length(target_scenario)){
  
    params_scenario[[i]] <- MizerParams(df_param,
                            interaction = theta, 
                            kappa = kappa,
                            kappa_ben = kappa_ben, 
                            kappa_alg = kappa_alg, 
                            w_pp_cutoff = w_pp_cutoff,
                            min_w_bb = min_w_bb, 
                            w_bb_cutoff = w_bb_cutoff, 
                            fleetDynamics = fleetDynamics, 
                            selectivity_params = selectivity_params,
                            catchability = target_scenario[[i]],
                            target = target_scenario[[i]])

    sim_scenario[[i]] <- project(params_scenario[[i]], 
                      effort = effort, 
                      dt = dt, 
                      fleetDynamics = fleetDynamics, 
                      management = fleetDynamics, 
                      multiFleet = multiFleet,
                      price = price_forward, 
                      cost = cost_forward, 
                      diet_steps = diet_steps, 
                      ke = ke,  
                      initial_effort = initial_effort_scenario,
                      scaling_price = scaling_price,
                      Blevel_management = Blevel_management,
                      initial_n = initial_n_scenario,
                      initial_n_pp = initial_n_pp_scenario, 
                      initial_n_bb = initial_n_bb_scenario)
  
    names(sim_scenario)[i]<-names(target_scenario)[i]
  
  } 
  
  return(sim_scenario)
}

sim_scenario<-runModel(target_scenario)

```

### Figure 4

```{r fig 4, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=10}

######## delete unfished as not used here 
sim_scenario$unfished<-NULL

######## order and rename variables 
# scenarios
a = c("statusQuo","noCompetition","fullCompetition","noBycatch","MoreUnderQuota","MoreTarget")
b = c("Status Quo","No Comp.","Max Comp.", "Bycatch","Under-utilised", "Valuable")

# fleets 
d = c("SET-SH","SET-US","SSG","SET-DS","SED")

# species  
spNames<-c("Lanternfish","Whiting","Squid","Perch","Mackerel","Redfish","Deep shark","Morwong","Flathead","Dories","Blue warehou","Orange roughy","Blue granadier","Silver warehou","Gemfish","Pink ling","Sawshark","Gummy shark","School shark")

# indicators 
e = c("profit", "yield", "effort", "Fref","Nref48", "Nref40","Nref20","slope", "biomassTarget","biomassSens","biomass")
f = c("Profit", "Yield", "Effort", "Active fleets","Spp above Bmey","Spp above Bmsy","Spp above Blim", "Slope","Biomass target", "Biomass sensitive","Biomass")

######## Fig 
# panel A - fleet-species interaction matrix for each scenario
plot_matrix = plotFleetMatrix(a,b,target_scenario)$plot_matrix 
plot_matrix <- plot_matrix + ggtitle("a)") + theme(plot.title = element_text(size = 11, face = "bold"))

# panel B top - effort dynamics for each fleet and scenarios 
col_values<-c("#b2182b", "#ef8a62", "#d1e5f0","#67a9cf", "#2166ac")
plot_effort = plotFleetEffort(a,b, sim_scenario, sim, col_values)

# panel B bottom - trends in main indicators (biomass of main target species, yields and profits at the fishery level)  
col_values<-c("#7b3294","#c2a5cf","#a6dba0","#008837")
plot_indiTrend<-indicatorsTrend(a,b, sim_scenario, sim, col_values) 
plot_all<-plot_indiTrend$plot_indiTrend

# resize and rearrange for publication
plot_matrix<-plot_matrix +theme_final+ theme(strip.text.x = element_text(face = "plain"), legend.key.size = unit(0.35, 'cm'), axis.text.x = element_text(size=8, hjust=0.95))

plot_effort<-plot_effort + theme_final+ theme(strip.text.x = element_text(face = "plain"), legend.key.size = unit(0.35, 'cm'), axis.title.x=element_blank(), axis.text.x=element_blank(),                 axis.ticks.x=element_blank())
plot_effort<-plot_effort+ ggtitle("b)") + theme(plot.title = element_text(size = 11, face = "bold"))

plot_all<-plot_all + theme_final+ theme(legend.key.size = unit(0.35, 'cm'), strip.text.x = element_blank())

# merge panel B top and bottom rows
fig4B<-plot_effort/plot_all

# Fig 4
pdf("/Users/camillan/Dropbox/Mizer-fleet_extension/plot/FD/Final/revision/Final_uploaded/Figure_4_flipped.pdf", height=7.03, width=9)
plot_matrix / fig4B + plot_layout (nrow = 2, heights = c(2.2,2.8))
dev.off()

# pdf("/Users/camillan/Dropbox/Mizer-fleet_extension/plot/FD/Final/revision/Final_uploaded/Figure_4.pdf", height=7.03, width=7.03)
# plot_matrix / fig4B + plot_layout (nrow = 2, heights = c(2.2,2.8))
# dev.off()

```

### Figure 5

```{r fig 5, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, fig.width=6, fig.height=6}

# avoid scenarios abbreviations for this fig 
b = c("Status Quo","No Competition","Max Competition", "Bycatch","Under-utilised", "Valuable")

######## Fig 
indi<-indicators(sim_scenario, management = TRUE)
df_plot = indi$df_plot

col_values3<-c("#7fbf7b","#af8dc3")
lolliIndicator = plotIndicators(a,b,df_plot, col_values3, scenarios = "all") 
lolliIndicator1 <- lolliIndicator$lolliIndicator1
lolliIndicator2 <- lolliIndicator$lolliIndicator2

lolliIndicator1 <- lolliIndicator1 + ggtitle("a)") + theme(plot.title = element_text(size = 11, face = "bold"))
lolliIndicator2 <- lolliIndicator2 + ggtitle("b)") + theme(plot.title = element_text(size = 11, face = "bold"))

# resize for publication 
lolliIndicator1<-lolliIndicator1 +theme_final + theme(strip.text.x = element_text(face = "plain"))
lolliIndicator2<-lolliIndicator2 +theme_final + theme(strip.text.x = element_text(face = "plain"))

# final Fig 5
pdf("/Users/camillan/Dropbox/Mizer-fleet_extension/plot/FD/Final/revision/Final_uploaded/Figure_5.pdf", height=7.08, width=7.08)
lolliIndicator1/lolliIndicator2 + plot_layout (nrow = 2, heights = c(2.1,0.9))
dev.off()

```



